---
title: "Characterization of Childhood Immunization Coverage in Kenya"
author: "David Apamo"
date: "`r Sys.Date()`"
output: word_document
---

# Introduction

Child immunization is a highly cost-effective program which fosters health equity for the entire population. According to the World Health Organization Immunization Agenda 2030, equity and coverage are strategic priority goals to ensuring that everyone, regardless of where they live, can access life-saving vaccines. Kenya has made notable progress in expanding its immunization programs over the years, resulting in increased vaccination coverage and a reduction in vaccine-preventable diseases. The government, in collaboration with both national and international partners and organizations e.g. KEMRI and WHO, has implemented various policies and strategies to enhance immunization access and equity all over the country. Despite these efforts, disparities are still present in immunization coverage. The disparities are driven by socioeconomic, maternal, geographic, child, and place of birth factors.

The disparities in immunization coverage can have profound implications for the overall health and well-being of children and can undermine the achievement of immunization targets set by the World Health Organization, as most children remain vulnerable to vaccine-preventable diseases, leading to increased morbidity and mortality rates. Addressing these inequities is crucial to achieving universal immunization coverage and improving the health outcomes of all children in Kenya.

# Study Objective

The main objective of this study was to assess the coverage of childhood immunization in Kenya among children aged 12 – 23 months based on socioeconomic, geographic, maternal, child, and place of birth characteristics.

```{r}
# Load packages
suppressMessages({
  library(haven)
  library(tidyverse)
  library(sjPlot)
  library(sjmisc)
  library(sjlabelled)
  library(gtsummary)
})
```

```{r}
# Import data
Child_immunisation <- read_dta("Child_immunisation.dta")
```


The data contains 3561 observations of 1340 variables.

```{r}
# View the first six observations
head(Child_immunisation)
```

```{r}
# Select the variables of interest
data <- Child_immunisation |> select(child_age, matage, matagebirth, BCG, DPT1,
                                     DPT2,DPT3, Polio_0, Polio_1, Polio_2, Polio_3,
                                     Measles_1, Measles_2, Pneumococcal_1, Pneumococcal_2,
                                     Pneumococcal_3, Rotavirus_1, Rotavirus_2, VitaminA_1, 
                                     Pentavalent_1, Pentavalent_2, Pentavalent_3, 
                                     Hepatitis_B1, Hepatitis_B2, Hepatitis_B3, HiB_1,
                                     HiB_2, HiB_3, full_vac, basic_vac, BCG_full, DPT_full,
                                     Polio_full, Measles_full, Pneumococcal_full,
                                     Rotavirus_full, VitaminA_full, Pentavalent_full,
                                     Hepatitis_B_full, HiB_full, birth_order, matagebirth3,
                                     motherhhd, delivery, sex, educ, wealth, religion,
                                     rural,union, county)
```

```{r}
# View the structure of the reduced data
glimpse(data)
```

We've selected 51 variables of interest, observations are still 3561.

# Descriptive characteristics 

We'll calculate the mean coverage of full immunization, as well as the mean coverage of individual vaccines at the National Level, in order to determine the overall childhood immunization coverage rates in Kenya.

### Vaccination coverage in Kenya at the National level (Mean coverage with 95% CI, by vaccine)

```{r}
# Tabulate the frequencies of full vaccination
table(data$full_vac)
```

Only 586 children out of 2975 received full vaccination. That is, children who received 1 of dose BCG, 3 doses of Polio, 3 doses of DPT, 2 doses of Measles, 2 doses of Rotavirus, 3 doses of Pentavalent, 3 doses of Pneumococcal, 3 doses of Hepatitis B, 3-dose HiB and 1 dose of Vitamin A vaccines.

```{r}
# Full vaccination mean
full_vac_mean <- mean(data$full_vac)
full_vac_mean
# CI for full_vac mean
CI <- prop.test(sum(data$full_vac), 3561, conf.level = 0.95)$conf.int
CI
```

Only 16.46% of the children were fully vaccinated. CI = [15.26, 17.72]

```{r}
# Basic vaccination mean
basic_vac_mean <- mean(data$basic_vac)
basic_vac_mean
# CI for basic vaccination
basic_vac_CI <- prop.test(sum(data$basic_vac), 3561, conf.level = 0.95)$conf.int
basic_vac_CI
```

Only 19.52% of the children received full basic vaccination, CI = [18.23, 20.87]. That is, children who received 1 dose of BCG, 3 doses of Polio, 3 doses of DPT and 2 doses of Measles vaccines.

```{r}
# Polio_full mean
Polio_full_mean <- mean(data$Polio_full)
Polio_full_mean
# CI for Polio_full vaccination
Polio_full_CI <- prop.test(sum(data$Polio_full), 3561, conf.level = 0.95)$conf.int
Polio_full_CI
```

75.344% received full Polio vaccination (Polio 1, 2 and 3). CI = [73.89, 76.75]

```{r}
# BCG mean (BCG was only given once at birth)
BCG_mean <- mean(data$BCG)
BCG_mean
# CI for BCG vaccination
BCG_CI <- prop.test(sum(data$BCG), 3561, conf.level = 0.95)$conf.int
BCG_CI
```

95.23% received the BCG vaccination given at birth. CI = [94.46, 95.89]

```{r}
# DPT_full mean
DPT_full_mean <- mean(data$DPT_full)
DPT_full_mean
# CI for DPT_full vaccination
DPT_full_CI <- prop.test(sum(data$DPT_full), 3561, conf.level = 0.95)$conf.int
DPT_full_CI
```

85.87% received full DPT vaccination (DPT 1, 2 and 3). CI = [84.68, 86.99]

```{r}
# Measles_full mean
Measles_full_mean <- mean(data$Measles_full)
Measles_full_mean
# CI for Measles_full_mean vaccination
Measles_full_mean_CI <- prop.test(sum(data$Measles_full), 3561, conf.level = 0.95)$conf.int
Measles_full_mean_CI
```

Only 24.57% of the children received full Measles vaccination (Measles 1 and 2). CI = [23.17, 26.03]

```{r}
# Pneumococcal_full mean
Pneumococcal_full_mean <- mean(data$Pneumococcal_full)
Pneumococcal_full_mean
# CI for Pneumococcal_full_mean vaccination
Pneumococcal_full_mean_CI <- prop.test(sum(data$Pneumococcal_full), 3561,
                                       conf.level = 0.95)$conf.int
Pneumococcal_full_mean_CI
```

87.78% received full Pneumococcal vaccination (Pneumococcal 1,2 and 3). CI = [86.65, 88.83]

```{r}
# Rotavirus_full mean
Rotavirus_full_mean <- mean(data$Rotavirus_full)
Rotavirus_full_mean
# CI for Rotavirus_full_mean vaccination
Rotavirus_full_mean_CI <- prop.test(sum(data$Rotavirus_full), 3561,
                                       conf.level = 0.95)$conf.int
Rotavirus_full_mean_CI
```

89.47% received full Rotavirus vaccination (Rotavirus 1,2 and 3). CI = [88.40, 90.45]

```{r}
# Vitamin_A full mean
VitaminA_full_mean <- mean(data$VitaminA_full)
VitaminA_full_mean
# CI for Vitamin_A_full_mean vaccination
VitaminA_full_mean_CI <- prop.test(sum(data$VitaminA_full), 3561,
                                       conf.level = 0.95)$conf.int
VitaminA_full_mean_CI
```

80.20% received full Vitamin A vaccination (Vitamin A1). CI = [78.85, 81.49]

```{r}
# Pentavalent_full mean
Pentavalent_full_mean <- mean(data$Pentavalent_full)
Pentavalent_full_mean
# CI for Pentavalent_full_mean vaccination
Pentavalent_full_mean_CI <- prop.test(sum(data$Pentavalent_full), 3561,
                                       conf.level = 0.95)$conf.int
Pentavalent_full_mean_CI
```

85.87% received full Pentavalent vaccination (Pentavalent 1, 2 and 3). CI = [84.68, 86.99]

```{r}
# Hepatitis_B_full mean
Hepatitis_B_full_mean <- mean(data$Hepatitis_B_full)
Hepatitis_B_full_mean
# CI for Hepatitis_B_full_mean vaccination
Hepatitis_B_full_mean_CI <- prop.test(sum(data$Hepatitis_B_full), 3561,
                                       conf.level = 0.95)$conf.int
Hepatitis_B_full_mean_CI
```

77.87% received full Hepatitis_B vaccination (Hepatitis_B 1,2 and 3). CI = [76.46, 79.22]

```{r}
# HiB_full mean
HiB_full_mean <- mean(data$HiB_full)
HiB_full_mean
# CI for HiB_full_mean vaccination
HiB_full_mean_CI <- prop.test(sum(data$HiB_full), 3561,
                                       conf.level = 0.95)$conf.int
HiB_full_mean_CI
```

85.87% received full HiB vaccination. CI = [84.68, 86.99]

## Individual vaccines

```{r}
# DPT 1 mean
DPT1_mean <- (sum(data$DPT1 ==1)/3561)
DPT1_mean
# CI for DPT 1 mean
DPT1_mean_CI <- prop.test(sum(data$DPT1 == 1), 3561,
                                       conf.level = 0.95)$conf.int
DPT1_mean_CI
```

94.78% of the children received the first dose of DPT vaccine. CI = [93.98, 95.47]

```{r}
# DPT 2 mean
DPT2_mean <- (sum(data$DPT2 ==1)/3561)
DPT2_mean
# CI for DPT 2 mean
DPT2_mean_CI <- prop.test(sum(data$DPT2 == 1), 3561,
                                       conf.level = 0.95)$conf.int
DPT2_mean_CI
```

90.99% received the second dose of DPT vaccine. CI = [89.99, 91.90]

```{r}
# DPT 3 mean
DPT3_mean <- (sum(data$DPT3 ==1)/3561)
DPT3_mean
# CI for DPT 3 mean
DPT3_mean_CI <- prop.test(sum(data$DPT3 == 1), 3561,
                                       conf.level = 0.95)$conf.int
DPT3_mean_CI
```

86.58% received the third dose of DPT vaccine. CI = [85.40, 87.67]

```{r}
# Polio 1 mean (We excluded the Polio vaccine given at birth)
Polio1_mean <- mean(data$Polio_1)
Polio1_mean
# CI for Polio_1 mean
Polio1_mean_CI <- prop.test(sum(data$Polio_1), 3561,
                                       conf.level = 0.95)$conf.int
Polio1_mean_CI
```

93.82% of the children received the first dose of Polio vaccine. CI = [92.97, 94.58]

```{r}
# Polio 2 mean
Polio2_mean <- mean(data$Polio_2)
Polio2_mean
# CI for Polio_2 mean
Polio2_mean_CI <- prop.test(sum(data$Polio_2), 3561,
                                       conf.level = 0.95)$conf.int
Polio2_mean_CI
```

90.96% received the second dose of Polio vaccine. CI = [89.96, 91.87]

```{r}
# Polio 3 mean
Polio3_mean <- mean(data$Polio_3)
Polio3_mean
# CI for Polio_3 mean
Polio3_mean_CI <- prop.test(sum(data$Polio_3), 3561,
                                       conf.level = 0.95)$conf.int
Polio3_mean_CI
```

76.41% received the third dose of Polio vaccine. CI = [74.97, 77.79]

```{r}
# Measles 1 mean
Measles1_mean <- (sum(data$Measles_1 == 1)/3561)
Measles1_mean
# CI for Measles1 mean
Measles1_mean_CI <- prop.test(sum(data$Measles_1 == 1), 3561,
                                       conf.level = 0.95)$conf.int
Measles1_mean_CI
```

83.43% of the children received the first dose of Measles vaccine. CI = [82.16, 84.63]

```{r}
# Measles 2 mean
Measles2_mean <- (sum(data$Measles_2 == 1)/3561)
Measles2_mean
# CI for Measles2 mean
Measles2_mean_CI <- prop.test(sum(data$Measles_2 == 1), 3561,
                                       conf.level = 0.95)$conf.int
Measles2_mean_CI
```

Only 25.78% of the children received the second dose of Measles vaccine. CI = [24.36, 27.26]

```{r}
# Rotavirus 1 mean
Rotavirus1_mean <- (sum(data$Rotavirus_1 == 1)/3561)
Rotavirus1_mean
# CI for Rotavirus1 mean
Rotavirus1_mean_CI <- prop.test(sum(data$Rotavirus_1 == 1), 3561,
                                       conf.level = 0.95)$conf.int
Rotavirus1_mean_CI
```

93.15% of the children received the first dose of Rotavirus vaccine. CI = [92.26, 93.95]

```{r}
# Rotavirus 2 mean
Rotavirus2_mean <- (sum(data$Rotavirus_2 == 1)/3561)
Rotavirus2_mean
# CI for Rotavirus2 mean
Rotavirus2_mean_CI <- prop.test(sum(data$Rotavirus_2 == 1), 3561,
                                       conf.level = 0.95)$conf.int
Rotavirus2_mean_CI

## Rotavirus 3 is not captured in the data
## Vitamin A was only one dose  and is captured in Vitamin_A_full mean and CI
```

90.23% received the second dose of Rotavirus vaccine. CI = [89.19, 91.17]

```{r}
# Pentavalent 1 mean
Pentavalent1_mean <- (sum(data$Pentavalent_1 == 1)/3561)
Pentavalent1_mean
# CI for Pentavalent_1 mean
Pentavalent1_mean_CI <- prop.test(sum(data$Pentavalent_1 == 1), 3561,
                                       conf.level = 0.95)$conf.int
Pentavalent1_mean_CI
```

94.78% of the children received the first dose of Pentavalent vaccine. CI = [93.98, 95.47]

```{r}
# Pentavalent 2 mean
Pentavalent2_mean <- (sum(data$Pentavalent_2 == 1)/3561)
Pentavalent2_mean
# CI for Pentavalent_2 mean
Pentavalent2_mean_CI <- prop.test(sum(data$Pentavalent_2 == 1), 3561,
                                       conf.level = 0.95)$conf.int
Pentavalent2_mean_CI
```

90.99% of the children received the second dose of Pentavalent vaccine. CI = [89.99, 91.90]

```{r}
# Pentavalent 3 mean
Pentavalent3_mean <- (sum(data$Pentavalent_3 == 1)/3561)
Pentavalent3_mean
# CI for Pentavalent_3 mean
Pentavalent3_mean_CI <- prop.test(sum(data$Pentavalent_3 == 1), 3561,
                                       conf.level = 0.95)$conf.int
Pentavalent3_mean_CI
```

86.58% of the children received the third dose of Pentavalent vaccine. CI = [85.40, 87.67]

```{r}
# Pneumococcal 1 mean
Pneumococcal_1_mean <- (sum(data$Pneumococcal_1 == 1)/ 3561)
Pneumococcal_1_mean
# CI for Pneumococcal_1 mean
Pneumococcal_1_mean_CI <- prop.test(sum(data$Pneumococcal_1 == 1), 3561,
                                       conf.level = 0.95)$conf.int
Pneumococcal_1_mean_CI
```

94.1% of the children received the first dose of Pneumococcal vaccine. CI = [93.27, 94.84]

```{r}
# Pneumococcal 2 mean
Pneumococcal_2_mean <- (sum(data$Pneumococcal_2 == 1)/ 3561)
Pneumococcal_2_mean
# CI for Pneumococcal_2 mean
Pneumococcal_2_mean_CI <- prop.test(sum(data$Pneumococcal_2 == 1), 3561,
                                       conf.level = 0.95)$conf.int
Pneumococcal_2_mean_CI
```

92.75% received the second dose of Pneumococcal vaccine. CI = [91.84, 93.57]

```{r}
# Pneumococcal 3 mean
Pneumococcal_3_mean <- (sum(data$Pneumococcal_3 == 1)/ 3561)
Pneumococcal_3_mean
# CI for Pneumococcal_3 mean
Pneumococcal_3_mean_CI <- prop.test(sum(data$Pneumococcal_3 == 1), 3561,
                                       conf.level = 0.95)$conf.int
Pneumococcal_3_mean_CI
```

88.54% received the third dose of Pneumococcal vaccine. CI = [87.44, 89.56]

```{r}
# Hepatitis B 1 mean
Hepatitis_B1_mean <- (sum(data$Hepatitis_B1 ==1)/3561)
Hepatitis_B1_mean
# CI for Hepatitis_B1 mean
Hepatitis_B1_mean_CI <- prop.test(sum(data$Hepatitis_B1 == 1), 3561,
                                       conf.level = 0.95)$conf.int
Hepatitis_B1_mean_CI
```

84.64% of the children received the first dose of Hepatitis B vaccine. CI = [83.40, 85.80]

```{r}
# Hepatitis B 2 mean
Hepatitis_B2_mean <- (sum(data$Hepatitis_B2 ==1)/3561)
Hepatitis_B2_mean
# CI for Hepatitis_B2 mean
Hepatitis_B2_mean_CI <- prop.test(sum(data$Hepatitis_B2 == 1), 3561,
                                       conf.level = 0.95)$conf.int
Hepatitis_B2_mean_CI
```

90.99% received the second dose of Hepatitis B vaccine. CI = [89.99, 91.90]

```{r}
# Hepatitis B 3 mean
Hepatitis_B3_mean <- (sum(data$Hepatitis_B3 ==1)/3561)
Hepatitis_B3_mean
# CI for Hepatitis_B3 mean
Hepatitis_B3_mean_CI <- prop.test(sum(data$Hepatitis_B3 == 1), 3561,
                                       conf.level = 0.95)$conf.int
Hepatitis_B3_mean_CI
```

86.58% received the third dose of Hepatitis B vaccine. CI = [85.40, 87.67]

```{r}
# HiB 1 mean
HiB1_mean <- (sum(data$HiB_1 == 1)/3561)
HiB1_mean
# CI for HiB1 mean
HiB1_mean_CI <- prop.test(sum(data$HiB_1 == 1), 3561,
                                       conf.level = 0.95)$conf.int
HiB1_mean_CI
```

94.78% of the children received the first dose of HiB vaccine. CI = [93.98, 95.47]

```{r}
# HiB 2 mean
HiB2_mean <- (sum(data$HiB_2 == 1)/3561)
HiB2_mean
# CI for HiB2 mean
HiB2_mean_CI <- prop.test(sum(data$HiB_2 == 1), 3561,
                                       conf.level = 0.95)$conf.int
HiB2_mean_CI
```

90.99% received the second dose of HiB vaccine. CI = [89.99, 91.90]

```{r}
# HiB 3 mean
HiB3_mean <- (sum(data$HiB_3 == 1)/3561)
HiB3_mean
# CI for HiB3 mean
HiB3_mean_CI <- prop.test(sum(data$HiB_3 == 1), 3561,
                                       conf.level = 0.95)$conf.int
HiB3_mean_CI
```

86.58% of the children received the third dose of HiB vaccine. CI = [85.40, 87.67]

```{r}
## Factor the variables of interest (socioeconomic, geographical, maternal, child and place of birth characteristics)

# Factor birth order variable
data[["birth_order"]] <- factor(data[["birth_order"]],
                                       labels = c("First","Second-Third",
                                                  "Fourth-Fifth","Six-up"), 
                                levels = c(1,2,3,4))

# Factor maternal age at birth(matagebirth3)
data[["matagebirth3"]] <- factor(data[["matagebirth3"]],
                                       labels = c("less than 20 years", "20-29 years", 
                                                  "30 years or more"), 
                                 levels = c(1,2,3))

# Factor mother head of household variable
data[["motherhhd"]] <- factor(data[["motherhhd"]],
                                       labels = c("Other","Mother"),
                                       levels = c(0,1))

# Factor the variable delivery (place of birth setting)
data[["delivery"]] <- factor(data[["delivery"]],
                                       labels = c("Home","Clinical Setting","Other"),
                                       levels = c(1,2,3))

# Factor sex of child variable
data[["sex"]] <- factor(data[["sex"]],
                                       labels = c("Male","Female"),
                                       levels = c(1,2))


# Factor maternal education variable
data[["educ"]] <- factor(data[["educ"]],
                                  labels = c("No education", "Primary", "Secondary", 
                                             "Higher"), levels = c(1,2,3,4))


# Factor wealth status variable
data[["wealth"]] <- factor(data[["wealth"]],
                                  labels = c("Poorest", "Poorer", "Middle", 
                                             "Richer", "Richest"), 
                           levels = c(1,2,3,4,5))


# Factor the variable religion
data[["religion"]] <- factor(data[["religion"]],
                                  labels = c("No religion", "Muslim", 
                                             "Christian", "Other"), 
                             levels = c(1,2,3,4))

# Factor the variable rural(place of residence) variable
data[["rural"]] <- factor(data[["rural"]],
                                  labels = c("Rural", "Urban"), levels = c(1,2))

# Factor the variable union(Partnership status)
data[["union"]] <- factor(data[["union"]],
                                  labels = c("Formerly", "Currently", "Never"),
                                  levels = c(1,2,3))

# Factor the variable county(region)
data[["county"]] <- factor(data[["county"]],
                                  labels = c("Mombasa", "Kwale", "Kilifi", "Tana River",
                                             "Lamu", "Taita taveta", "Garissa", "Wajir",
                                             "Mandera", "Marsabit", "Isiolo", "Meru",
                                             "Taraka-nithi", "Embu", "Kitui","Machakos",
                                             "Makueni", "Nyandarua", "Nyeri",
                                             "Kirinyaga", "Murang'a", "Kiambu",
                                             "Turkana", "West-Pokot", "Samburu",
                                             "Trans-nzoia", "Uasian-gishu",
                                             "Elgeyo-marakwet", "Nandi", "Baringo",
                                             "Laikipia", "Nakuru", "Narok", "Kajiado",
                                             "Kericho", "Bomet", "Kakamega", "Vihiga",
                                             "Bungoma", "Busia", "Siaya", "Kisumu",
                                             "Homa-bay", "Migori", "Kisii", "Nyamira",
                                             "Nairobi"),
                                  levels = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17
                                            ,18,19,20,21,22,23,24,25,26,27,28,29,30
                                            ,31,32,33,34,35,36,37,38,39,40,41,42,43,
                                             44,45,46,47))


```

# Data Visualization

We want to visualize the socioeconomic, geographical, maternal, child and place of birth characteristics based on full vaccination. The plots are only for full vaccinated children.

```{r}
# First filter the data for children who were fully vaccinated
full_vac_data <- data |> filter(full_vac == 1)
```

```{r}
# Full vaccination Vs Birth order
ggplot(full_vac_data, aes(x = birth_order, fill = birth_order)) + 
  geom_bar() + 
  labs(title = "Birth Order", 
       x = NULL, y = "Frequency", 
       fill = "Birth Order") +
  coord_flip()
```

The most vaccinated children were second or third born. Quite a number of firstborns also received full vaccination.

```{r}
# Full vaccination Vs Maternal age at birth
ggplot(full_vac_data, aes(x = matagebirth3, fill = matagebirth3)) + 
  geom_bar() + 
  labs(title = "Maternal Age at Birth", 
       x = NULL, y = "Frequency", 
       fill = "Maternal Age at Birth") + 
  coord_flip()
```

Most of the fully vaccinated children were for mothers aged 20 to 29 years, followed by mothers aged 30 years and above.

```{r}
# Full vaccination Vs Household Head
ggplot(full_vac_data, aes(x = motherhhd, fill = motherhhd)) + 
  geom_bar() + 
  labs(title = "Head of Household", 
       x = NULL, 
       y = "Frequency", 
       fill = "Head of Household") + 
  coord_flip()
```

Most of the children who received full vaccination came from households which are not headed by the mother.

```{r}
# Full vaccination Vs Delivery(place of birth setting)
ggplot(full_vac_data, aes(x = delivery, fill = delivery)) + 
  geom_bar() + 
  labs(title = "Place of Birth Setting", 
       x = NULL, y = "Frequency", 
       fill = "Place of Birth Setting") + 
  coord_flip()
```

Most of the children who were fully vaccinated were born in a clinical setting.

```{r}
# Full vaccination Vs sex of child
ggplot(full_vac_data, aes(x = sex, fill = sex)) + 
  geom_bar() + 
  labs(title = "Sex of Child", 
       x = NULL, y = "Frequency", 
       fill = "Sex") + 
  coord_flip()
```

Male babies were more than female babies, that is for those who were fully vaccinated. The difference is however minimal.

```{r}
# Full vaccination Vs maternal education
ggplot(full_vac_data, aes(x = educ, fill = educ)) + 
  geom_bar() + 
  labs(title = "Maternal Education", 
       x = NULL, y = "Frequency", 
       fill = "Maternal Education") + 
  coord_flip()
```

Most of the fully vaccinated children were for mothers with secondary education, followed by primary then tertiary education. Few mothers with no formal education also got their children vaccinated.

```{r}
# Full vaccination Vs wealth status
ggplot(full_vac_data, aes(x = wealth, fill = wealth)) + 
  geom_bar() + 
  labs(title = "Wealth Status", 
       x = NULL, y = "Frequency", 
       fill = "Wealth Status") + 
  coord_flip()
```

Most children who were fully vaccinated came from richer households.

```{r}
# Full vaccination Vs Religion
ggplot(full_vac_data, aes(x = religion, fill = religion)) + 
  geom_bar() + 
  labs(title = "Religion", x = NULL, 
       y = "Frequency", fill = "Religion") + 
  coord_flip()
```

Most of the fully vaccinated children were from christian homes.

```{r}
# Full vaccination Vs Place of residence(rural/urban)
ggplot(full_vac_data, aes(x = rural, fill = rural)) + 
  geom_bar() + 
  labs(title = "Residential Place", 
       x = NULL, y = "Frequency", 
       fill = "Residential Place") + 
  coord_flip()

```

Most of the fully vaccinated children were from rural home settings.

```{r}
# Full vaccination Vs union/partnership
ggplot(full_vac_data, aes(x = union, fill = union)) + 
  geom_bar() + 
  labs(title = "Partnership Status", 
       x = NULL, y = "Frequency", 
       fill = "Partnership Status") + 
  coord_flip()
```

Children whose mothers are currently married were the highest with regards to full vaccination.

```{r}
# Full vaccination Vs County/region
ggplot(full_vac_data, aes(x = county, fill = county)) + 
  geom_bar() + 
  labs(title = "County/Region", x = NULL, 
       y = "Frequency", fill = "County") + 
  coord_flip()
```

Most of the fully vaccinated children were from Migori county, followed by Kakamega, Bungoma, Nakuru and Kwale. We noted that Garissa county only had 1 fully immunized child, Mandera had 2 and Wajir had 3.


# Measures of Association between full Immunization and Socioeconomic, Geographical, Maternal, Child and Place of birth characteristics.

We want to test the hypothesis that;

* Ho: There is no association between full vaccination and the socioeconomic, geographical, maternal, child and place of birth characteristics.
* Ha: There is association between full vaccination and the socioeconomic, geographical, maternal, child and place of birth characteristics.

## 1. Socioeconomic Characteristics.

```{r}
## 1. Wealth status

# Contingency table
Table_1 <- xtabs(~ full_vac + wealth, data = data)
Table_1
```

```{r}
# Chi-square test of independence
chi_1 <- chisq.test(Table_1)
chi_1
```

The p-value < 0.05, implying that there's an association between full vaccination and wealth status.

```{r}
## 2. Union/partnership status

# Contingency table
Table_2 <- xtabs(~ full_vac + union, data = data)
Table_2
```

```{r}
# Chi-square test of independence
chi_2 <- chisq.test(Table_2)
chi_2
```

The p-value > 0.05, implying that there's no association between full vaccination and union/partnership status.

```{r}
## 3. Religion

# Contingency table
Table_3 <- xtabs(~ full_vac + religion, data = data)
Table_3
```

```{r}
# Chi-square test of independence
chi_3 <- chisq.test(Table_3)
chi_3
```

The p-value < 0.05, implying that there's an association between full vaccination and religion.

## 2.Geographical Characteristics

```{r}
## 4. Place of residence

# Contingency table
Table_4 <- xtabs(~ full_vac + rural, data = data)
Table_4
```

```{r}
# Chi-square test of independence
chi_4 <- chisq.test(Table_4)
chi_4
```

The p-value < 0.05, implying that there's an association between full vaccination and place of residence.

```{r}
## 5. County

# Contingency table
Table_5 <- xtabs(~ full_vac + county, data = data)
Table_5
```

```{r}
# Chi-square test of independence
chi_5 <- chisq.test(Table_5)
chi_5
```

The p-value < 0.05, implying that there's an association between full vaccination and county/region.

## 3. Maternal Characteristics

```{r}
## 6. Maternal Education

# Contingency table
Table_6 <- xtabs(~ full_vac + educ, data = data)
Table_6
```

```{r}
# Chi-square test of independence
chi_6 <- chisq.test(Table_6)
chi_6
```

The p-value < 0.05, implying that there's an association between full vaccination and maternal education.

```{r}
## 7. Mother household head

# Contingency table
Table_7 <- xtabs(~ full_vac + motherhhd, data = data)
Table_7
```

```{r}
# Chi-square test
chi_7 <- chisq.test(Table_7)
chi_7
```

The p-value > 0.05, implying that there's no association between full vaccination and the mother being household head.

```{r}
## 8. Maternal Age at Birth

# Contingency table
Table_8 <- xtabs(~ full_vac + matagebirth3, data = data)
Table_8
```

```{r}
# Chi-square test of independence
chi_8 <- chisq.test(Table_8)
chi_8
```

The p-value < 0.05, implying that there's an association between full vaccination and the maternal age at birth.

## 4. Child Characteristics

```{r}
## 9. Birth Order

# Contingency table
Table_9 <- xtabs(~ full_vac + birth_order, data = data)
Table_9
```

```{r}
# Chi-square test of independence
chi_9 <- chisq.test(Table_9)
chi_9
```

The p-value < 0.05, implying that there's an association between full vaccination and the birth order.

```{r}
## 10. Sex of child

# Contingency table
Table_10 <- xtabs(~ full_vac + sex, data = data)
Table_10
```

```{r}
# Chi-square test of independence
chi_10 <- chisq.test(Table_10)
chi_10
```

The p-value > 0.05, implying that there's no association between full vaccination and the sex of the child.

## 5. Place of Birth Characteristics

```{r}
## 11. Delivery (Place of birth setting)

# Contingency table
Table_11 <- xtabs(~ full_vac + delivery, data = data)
Table_11
```

```{r}
# Chi-square test of independence
chi_11 <- chisq.test(Table_11)
chi_11
```
 
The p-value < 0.05, implying that there's an association between full vaccination and the delivery/place of birth setting. There's a warning due to the very small number of observations in the last cell of the contingency table.


# Logistic Regression Model

We'll fit a Multiple Logistic Regression Model so that we can describe the relationship between full vaccination and the socioeconomic, geographical, maternal, child and place of birth characteristics. We'll only include variables that have shown significant associations with full vaccination from the chi-square tests of independence.

```{r}
# Fit the model
Logistic_Model <- glm(full_vac ~ wealth + religion + rural + educ +
                        matagebirth3 + birth_order + delivery + county, data = data,
                      family = "binomial")

```

```{r}
# Model summary
summary(Logistic_Model)
```


# Create Table for the Logistic Regression Model Results

I'll use the tab_model() function from sjPlot package to create a table for the Multiple Logistic Regression model. The function will convert the log-odds into odds ratios, which are easy to interpret.

```{r}
# Create table for the model
tab_model(Logistic_Model)
```

## Socioeconomic characteristics

* Children born in richer families had 67% higher odds of full immunization compared to children born in poorest families. 
* Children born in richest households had 78% higher odds of full immunization compared to children born in poorest households.

## Maternal Characteristics

* The odds for children of mothers with primary education being fully immunized were 2.02 times the odds for children whose mothers had no education.
* The odds for children of mothers with secondary education being fully immunized were 1.95 times the odds for children whose mothers had no education.
* The odds for children of mothers with higher (college or university) education being fully immunized were 1.83 times the odds for children whose mothers had no education.
* Children of mothers aged 20-29 years had 68% higher odds of full immunization compared to children of mothers aged below 20 years. 
* Children of mothers aged 30 years and above had 89% higher odds of full immunization compared to children of mothers aged below 20 years. 

## Child characteristics

* Children who were sixth born or higher had 57% lower odds of full immunization.

## Geographical Characteristics

Mombasa county was our reference level, based on how the data was coded.

* Children in Garissa county had 88.42% lower odds of full immunization compared to children in Mombasa county. 
* Children in Mandera county had 84.71% lower odds of full immunization in comparison to children in Mombasa county. 
* Children in West Pokot county had 64.8% lower odds of full immunization compared to children in Mombasa county. 
* Children in Samburu county had 71.38% lower odds of full immunization compared to children in Mombasa county. 
* Children in Narok county had 62.82% lower odds of full immunization compared to children in Mombasa county. 

The Logistic Regression model though, had a lower adjusted R2 because most of the independent variables in the model were not statistically significant.

```{r}
# Save the data as a CSV file
write.csv(data, file = "Child_Immunization_data.csv")
```

# References

Government of Kenya Ministry of Health. National Policy Guidelines on Immunization 2013 [Internet]. 2014 [cited 30 Apr 2020]. http://guidelines.health.go.ke:8000/media/Immunization_Policy_Guidline.pdf

Kenya National Bureau of Statistics. The DHS Program - Kenya: Standard DHS, 2014 [Internet].
https://doi.org/10.1101/2021.02.14.21251721; this version posted February 16, 2021

WHO | Global Vaccine Action Plan 2011-2020 [Internet]. 2012 [cited 5 Nov 2018]. https://www.who.int/immunization/global_vaccine_action_plan/GVAP_doc_2011_2020/en/

WHO. WHO | Explorations of inequality: Childhood immunization [Internet]. 2018 [cited 18 Sep 2019]. Available: https://www.who.int/gho/health_equity/report_2018_immunization/en/
